Statistical Inference for inter-arrival times of extreme events in bursty time series

### 2020-07-27 article_springer/article_springer.Rmd
Move Hill plot

Suppose the "correct" threshold $\ell_0$ was known, in the sense that 
thresholding at $\ell_0$ would result precisely in an underlying MRP 
$\{(W_i, J_i): i \in I\}$ as defined in Section \ref{sec:CTRE}. 
In this case, we could apply any tail (and scale) estimator such as e.g.
Hill's estimator to the $W_i$ and thus infer the distribution of 
inter-exceedance times.

@@

Due to the form of the density function of a Mittag-Leffler distribution there exists no closed form for the Maximum Likelihood estimator (MLE). In the R Package `MittagLeffleR` [@MittagLeffleR], Maximum Likelihood estimation is implemented via numerical optimization. The MLE slightly outperforms the log-Moment estimator regarding bias and RMSE for big enough sample sizes, but is extremely computationally intensive. Figure \ref{fig:MSE} shows that both estimators perform well, even for small sample sizes. For smaller tails both estimators show an increasing RMSE for the scale estimation due to an increasing variance. This results from the fact that in case of very small tails, single very large values can occur.

@@

hill_plot_values<-data.frame(x=numeric(),y=numeric(),rep=numeric())
for  (m in 1:100){
  data_ml_2<-rml(n=10000,tail=0.8,scale=1)
  df<-hill_estim_range(k_min=30,k_max=200,data=data_ml_2,true_tail=0.8)
  df<-df %>% mutate(rep=m)
  hill_plot_values<-bind_rows(df,hill_plot_values)
}
mean_n1000<-hill_plot_values %>% group_by(x) %>% summarize(mean=mean(y))

p2<-hill_plot_values %>% ggplot() + geom_line(aes(x=x,y=y,group=rep), alpha=0.1) +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(aes(x=x, y=mean),colour="black",data=mean_n1000) + 
  ylab("Estimated tail") +
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
  xlab("Order Statistics")+
  ylab("Hill-Estimator of Tail Index")

@@

>>>>

Suppose the "correct" threshold $\ell_0$ was known, in the sense that 
thresholding at $\ell_0$ would result precisely in an underlying MRP 
$\{(W_i, J_i): i \in I\}$ as defined in Section \ref{sec:CTRE}. 
In this case, we could apply any tail (and scale) estimator such as e.g.
Hill's estimator [@hill1975simple]
to the $W_i$ and thus infer the distribution of 
inter-exceedance times.

@@

For high thresholds where few IETs are left, estimating a tail parameter
by fitting a MLD has advantages over e.g. the Hill estimator: 
since the underlying distribution of IETs *is* a MLD, 
the fits are more accurate. 
Figure \ref{fig:Hillplots} shows Hill plots for 200 IETs, for two Mittag-Leffler
datasets of sizes 200 and 10000 (and an irrelevant distribution of magnitudes).
Clearly the variance is too high to give useful estimates.
For corresponding fits based on the assumption of a Mittag-Leffler distribution,
see Figure \ref{Fig:TailSimu}, top row, yielding more reliable predictions 
(more discussion further below).

@@

hill_plot_values<-data.frame(x=numeric(),y=numeric(),rep=numeric())
for  (m in 1:100){
  data_ml_2<-rml(n=10000,tail=0.8,scale=1)
  df<-hill_estim_range(k_min=30,k_max=200,data=data_ml_2,true_tail=0.8)
  df<-df %>% mutate(rep=m)
  hill_plot_values<-bind_rows(df,hill_plot_values)
}
mean_n10000<-hill_plot_values %>% group_by(x) %>% summarize(mean=mean(y))

p2<-hill_plot_values %>% ggplot() + geom_line(aes(x=x,y=y,group=rep), alpha=0.1) +
  coord_cartesian(ylim = c(0.5, 1.5)) +
  geom_line(aes(x=x, y=mean),colour="black",data=mean_n10000) + 
  ylab("Estimated tail") +
  geom_hline(yintercept=0.8, linetype="dashed", color = "red")+
  xlab("Order Statistics")+
  ylab("Hill-Estimator of Tail Index")

@@

The Maximum Likelihood Estimator (MLE) for the MLD is not straightforward 
to implement, since the MLD density admits no closed form analytical expression.
In the R Package `MittagLeffleR` [@MittagLeffleR], MLE is implemented via numerical optimization. The MLE slightly outperforms the log-Moment estimator regarding bias and RMSE for big enough sample sizes, but is computationally very intensive. Figure \ref{fig:MSE} shows that both estimators perform well, even for small sample sizes. For smaller tails both estimators show an increasing RMSE for the scale estimation due to an increasing variance. This results from the fact that in case of very small tails, single very large values can occur.
